Discrete Ordinates

### 2017-01-09 docprep/draft1.tex
Added a sequence diagram and added a section on the raytracer math

>>>>

\subsection{Raytracer}
Discrete ordinate solutions greatly benefit from the presence of a raytracing algorithm to provide the uncollided flux to the solver.

The raytracer was implemented as...

When the solver is changed from isotropic to anisotropic, the raytracer must be able to produce a Pn expansion needed by the anisotropic solver. We defined the spherical harmonic function as

\begin{equation}
    Y_l^m(\theta, \phi)= 
\begin{cases}
\sqrt{2} K_l^m \cos(m \phi) P_l^m(\cos(\theta)),      & m > 0 \\
K_l^0 P_l^0(\cos(\theta)),              & m = 0  \\
\sqrt{2} K_l^m \sin(-m \phi) P_l^{-m}(\cos(\theta)),           & m < 0
\end{cases}
\end{equation}

where $K_l^m$ is defined as

\begin{equation}
K_l^m = \sqrt{\frac{2l+1}{4 \pi} \frac{(l-|m|)!}{(l+|m|)!}}
\end{equation}

and the associated Legendre functions are computed by the recursion

\begin{equation}
\begin{aligned}
P_0^0(x) &= 1 \\
P_m^m(x) &= (2m-1)!!(1-x^2)^{m/2} \\
P_{m+1}^m(x) &= x(2m+1)P_m^m(x) \\
P_l^m(x) &= \frac{x(2l-1)}{l-m} P_{l-1}^m (x) - \frac{l+m-1}{l-m} P_{l-2}^m (x)
\end{aligned}
\end{equation}

The spherical harmonic expansion is used to approximate a function $f$ by

\begin{equation}
f(\theta, \phi) = \sum_{l=0}^\infty \sum_{m=-l}^l y_l^m(\theta, \phi) f_l^m
\end{equation}

which is exact when the expansion continues to infinity. However, it is trunctated at L yielding

\begin{equation}
f(\theta, \phi) \approx \sum_{l=0}^L \sum_{m=-l}^l y_l^m(\theta, \phi) f_l^m
\end{equation}

However, the value of each $f_l^m$ must be computed before the approximation can be used.

These values are computed by

\begin{equation}\label{eq:flm}
f_l^m = \int_{4 \pi} y_l^m(\theta, \phi) f(\theta, \phi) d\hat{\Omega}
\end{equation}

In the case of the raytracer solution, since \textit{only} the uncollided source is considered, the direction at any point is always precisely known, the particles are traveling away from the source point. Therefore, the function to be approximated is a Dirac delta function in the two dimensional surface of the unit sphere. Equation~\ref{eq:flm} can be reduced to

\begin{equation}
f_l^m = \int_{4 \pi} y_l^m(\theta, \phi)f(\theta_0, \phi_0) d\hat{\Omega}
\end{equation}

The property of the Dirac delta function

\begin{equation}
\int_{-\infty}^{\infty} \delta(x_0) f(x) dx = f(x_0)
\end{equation}

can now be used which reduces Equation~\ref{eq:flm} to

\begin{equation}
f_l^m = y_l^m(\theta_0, \phi_0)
\end{equation}

The moments are related to the angular flux by

\begin{equation}
\phi_l^m(\vec{r}, E) = \int Y_l^m(\hat{\Omega} ') \psi(\vec{r}, \hat{\Omega}', E) d\hat{\Omega}'
\end{equation}

which becomes

\begin{equation}
\phi_l^m(\vec{r}, E) = \Psi Y_l^m(\hat{\Omega}_0)
\end{equation}

where $\Psi$ is equal to the magnitude of $\psi_{unc}$ in the $\hat{\Omega}_0$ direction.
